---
title: "Parliament Data"
date: "AT 2023"
output: html_document
---

```{r setup, include=FALSE}

knitr::opts_chunk$set(echo = FALSE) 
```

## Part One: Retrieving Primary Data from UK Parliament API

### Oral and Written Questions

First, we retrieve the primary data from Parliament's API, including oral and written questions. Since the api is constantly updated with the latest questions asked, we set an end date of 31/12/2023 to keep results consistent through different runs.

By running the request in the API development hub, there are a total of 30878 entries of oral questions and 490441 entries of written questions. We set the take parameter to the maximum limit of 100 and loop through the entries by increasing the skip parameter.

The oral and written question APIs differ slightly in field names and in field content. In each oral question asked, there is a field detailing the information of the MP who asked the question, however, in written questions, this field is empty. To keep the the data consistent, we will only keep the MP Id for each question for now, and retrieve the MP information in the next step using the MP API. To reduce data size, we will also only keep the following columns: Id, QuestionText, TabledWhen, AnsweringBodyId, AskingMemberId, AnsweringBody for both oral and written questions.
```{r retrieve_primary_data, eval=FALSE}

library(httr)
library(jsonlite)
library(tidyverse)

# set the base url and end date
end_date <- "2023-12-31"
base_url_oral <- "https://oralquestionsandmotions-api.parliament.uk/oralquestions/list"
take <- 100
num_response <- 30878
num_request <- ceiling(num_response / take) # number of times to loop through the request

# set the tibble to store all the results from each run
oral_tibble <- tibble()

# loop through request to build the oral tibble
for (i in 0:(num_request - 1)) {
  skip <- i * take
  query <- sprintf("%s?parameters.answeringDateEnd=%s&parameters.skip=%s&parameters.take=%s", base_url_oral, end_date, skip, take)
  r <- GET(query)
  # check if the query is successful, print the parameters for unsuccessful queries
  if (status_code(r) != 200) {
    print("Error")
    print(skip)
    break
  }
  r_json <- fromJSON(rawToChar(r$content))
  r_tibble <- as_tibble(r_json$Response) |> 
    select(Id, QuestionText, TabledWhen, AnsweringBodyId, AskingMemberId, AnsweringBody) # select key rows
  oral_tibble <- oral_tibble |> bind_rows(r_tibble) # append to overall tibble
  Sys.sleep(1) # sleep for 1 second to avoid rate limit
}

# base url for written questions
base_url_written <- "https://questions-statements-api.parliament.uk/api/writtenquestions/questions"
written_take <- 10000
written_response <- 490441 # by observing the API response
written_request <- ceiling(written_response / written_take) # number of times to loop through the request

# set the tibble to store all the results from each run
written_tibble <- tibble()

# suppress scientific notation
options(scipen = 999)

# loop through request to build the written tibble
for (i in 0:(written_request - 1)) {
  skip <- i * written_take
  query <- sprintf("%s?tabledWhenTo=%s&skip=%s&take=%s", base_url_written, end_date, skip, written_take)
  print(query)
  r <- GET(query)
  # check if the query is successful, print the parameters for unsuccessful queries
  if (status_code(r) != 200) {
    print("Error")
    print(skip)
    break
  }
  r_json <- fromJSON(rawToChar(r$content))
  r_tibble <- as_tibble(r_json$results$value) |> 
    select(id, questionText, dateTabled, answeringBodyId, answeringBodyName) # select key rows
  written_tibble <- written_tibble |> bind_rows(r_tibble) # append to overall tibble
  Sys.sleep(5) # sleep for 5 seconds to avoid rate limit
}

# save both tibbles in .rds formats
saveRDS(oral_tibble, "data/oral_tibble.rds")
saveRDS(written_tibble, "data/written_tibble.rds")

```


```{r convert_json_to_tibble}

```

```{r read_json_files}
## Appendix: All code in this assignment

# this chunk generates the complete code appendix. 
# eval=FALSE tells R not to run (``evaluate'') the code here (it was already run before).
```
