---
title: "Parliament Data"
date: "AT 2023"
output: html_document
---

```{r setup, include=FALSE}

knitr::opts_chunk$set(echo = FALSE) 
```

## Part One: Retrieving Primary Data from UK Parliament API

### Oral and Written Questions

First, we retrieve the primary data from Parliament's API, including oral and written questions. Since the api is constantly updated with the latest questions asked, we set an end date of 31/12/2023 to keep results consistent through different runs.

By running the request in the API development hub, there are a total of 30878 entries of oral questions and 490441 entries of written questions. For oral questions, we set the take parameter to the maximum limit of 100 and loop through the entries by increasing the skip parameter. For written questions, there is no stated limit of the take parameter but through experimentation, setting the take parameter to 10000 strikes the balance between stability and speed. However, due to the large number of requests for written questions, the response sometimes unpredictably returns an error. To resolve this issue, we check the status code of each response, when it returns of a status code that is not 200, the programme will break from the current loop and print the request URL and its parameters that trigger the error. We can then manually adjust the skip parameter to continue the loop from the last successful request.

The oral and written question APIs differ slightly in field names and in field content. In each oral question asked, there is a field detailing the information of the MP who asked the question, however, in written questions, this field is empty. To keep the the data consistent, we will only keep the MP Id for each question for now, and retrieve the MP information in the next step using the MP API. To reduce data size, we will also only keep the following columns: Id, QuestionText, TabledWhen, AnsweringBodyId, AskingMemberId, AnsweringBody for both oral and written questions.

After retrieving the final tibbles, we will save the to .rds files for future use.

```{r retrieve_primary_data, eval=FALSE}

library(httr)
library(jsonlite)
library(tidyverse)

# set the base URL and end date
end_date <- "2023-12-31"
base_url_oral <- "https://oralquestionsandmotions-api.parliament.uk/oralquestions/list"
take <- 100
num_response <- 30878
num_request <- ceiling(num_response / take) # number of times to loop through the request

# set the tibble to store all the results from each run
oral_tibble <- tibble()

# loop through request to build the oral tibble
for (i in 0:(num_request - 1)) {
  skip <- i * take
  query <- sprintf("%s?parameters.answeringDateEnd=%s&parameters.skip=%s&parameters.take=%s", base_url_oral, end_date, skip, take)
  r <- GET(query)
  # check if the query is successful, print the parameters for unsuccessful queries
  if (status_code(r) != 200) {
    print("Error")
    print(skip)
    break
  }
  r_json <- fromJSON(rawToChar(r$content))
  r_tibble <- as_tibble(r_json$Response) |> 
    select(Id, QuestionText, TabledWhen, AnsweringBodyId, AskingMemberId, AnsweringBody) # select key rows
  oral_tibble <- oral_tibble |> bind_rows(r_tibble) # append to overall tibble
  Sys.sleep(1) # sleep for 1 second to avoid rate limit
}

# base URL for written questions
base_url_written <- "https://questions-statements-api.parliament.uk/api/writtenquestions/questions"
written_take <- 10000
written_response <- 490441 # by observing the API response
written_request <- ceiling(written_response / written_take) # number of times to loop through the request

# set the tibble to store all the results from each run
written_tibble <- tibble()

# suppress scientific notation
options(scipen = 999)

# loop through request to build the written tibble
for (i in 0:(written_request - 1)) {
  skip <- i * written_take
  query <- sprintf("%s?tabledWhenTo=%s&skip=%s&take=%s", base_url_written, end_date, skip, written_take)
  print(query)
  r <- GET(query)
  # check if the query is successful, print the parameters for unsuccessful queries
  if (status_code(r) != 200) {
    print("Error")
    print(skip)
    break
  }
  r_json <- fromJSON(rawToChar(r$content))
  r_tibble <- as_tibble(r_json$results$value) |> 
    select(id, questionText, dateTabled, answeringBodyId, answeringBodyName, askingMemberId) # select key rows
  written_tibble <- written_tibble |> bind_rows(r_tibble) # append to overall tibble
  print(i)
  Sys.sleep(5) # sleep for 5 seconds to avoid rate limit
}

# save both tibbles in .rds formats
saveRDS(oral_tibble, "data/oral_tibble.rds")
saveRDS(written_tibble, "data/written_tibble.rds")

```

### Filtering and Classifying Questions

Now we have the oral and written questions, we need to differentiate questions that are about economic issues and questions about health and welfare issues. First, we merge oral and written tibbles and add one column to store if each question is oral or written

There can be many strategies when it comes to filter the questions that are useful for our purpose. The most straightforward way is to filter the questions based on the answering bodies they are addressed to. However, dividing the questions purely based on the answering body proves to be less effective. This is largely because there are a total of 52 answering bodies and many of them do not have a clear focus on either economic or health and welfare issues. However, there are departments that are dedicated towards economic issues and health and welfare issues. They are:

|                                        |                                                            |
|-----------------------------|-------------------------------------------|
| Economics related Departments          | Treasury, HM Treasury                                      |
| Health and Welfare related Departments | Department of Health and Social Care, Department of Health |

Below are the step by step instructions on how the questions will be divided to two categories:

1.  Add one column to the tibble called "Issue Type" (Economic/Health and Welfare)
2.  Reference the table above, mark the questions addressed to the corresponding answering bodies
3.  For the questions that are not addressed to the answering bodies above, we will use the question text to determine which category the question belongs to. We will create a list of keywords that are related to economic issues and health and welfare issues. If the question text contains any of the keywords, we will mark the question as economic or health and welfare accordingly. The keywords are generated by ChatGPT 4 and double checked before being used.
4.  If the question text does not contain any of the keywords, we will mark the question as "Other". (this process may take longer to run)
5.  For text questions with words from both keywords list, we will mark the questions as "Pending" and duplicate them (one makred as Health and Welfare, one makred as Economic).

```{r filtering_and_classifying_questions}

# read the two tibbles from the .rds files
oral_tibble <- readRDS("data/oral_tibble.rds")
written_tibble <- readRDS("data/written_tibble.rds")

# merge two tibbles and add a column to indicate if the question is oral or written
oral_tibble <- oral_tibble |> mutate(QuestionType = "oral") 
written_tibble <- written_tibble |> mutate(QuestionType = "written")
# update the column names of written tibble to match oral tibble
colnames(written_tibble) <- c("Id", "QuestionText", "TabledWhen", "AnsweringBodyId", "AnsweringBody", "AskingMemberId", "QuestionType" )
# merge the two tibbles
questions_tibble <- oral_tibble |> bind_rows(written_tibble)

# Step one: add one column to the tibble called "Issue Type" (Economic/Health and Welfare)
questions_tibble <- questions_tibble |> mutate(IssueType = NA)

# Step two: Mark Economics and Health and Welfare related questions based on the answering bodies
# Economics related Departments: Treasury, HM Treasury, Department for Business and Trade
# Health and Welfare related Departments: Department of Health and Social Care, Department of Health
questions_tibble <- questions_tibble |> 
  mutate(IssueType = case_when(
  AnsweringBody == "Treasury" ~ "Economic",
  AnsweringBody == "HM Treasury" ~ "Economic",
  AnsweringBody == "Department of Health and Social Care" ~ "Health and Welfare",
  AnsweringBody == "Department of Health" ~ "Health and Welfare"
 ))

# Step three and four: keyword based classification

economic_keywords <- c(
  "economy", "inflation", "gdp", "unemployment", "fiscal policy", 
  "budget", "monetary policy", "investment", "taxation", 
  "debt", "financial markets", "economic growth", "currency", 
  "interest rates", "stock market", "deficit", "consumer spending", 
  "business regulation", "exchange rate", "economic development", 
  "banking", "financial crisis", "trade deficit", "inflation rate", 
  "economic policy", "entrepreneurship", "business growth", "market trends", 
  "economic stability", "investment strategy", "economic forecast", 
  "financial planning", "economic recovery", "economic reform",
  "housing market", "exports", "imports", "economic impact", "fiscal deficit",
  "labour market", "economic inequality", "economic sustainability", 
  "economic integration", "corporate tax", "economic diversification", 
  "economic sanctions", "economic indicators", "economic resilience"
)

health_welfare_keywords <- c(
  "nhs", "healthcare", "social care", "mental health", 
  "medicine", "public health", "insurance", "hospital", "disability", 
  "elderly care", "child welfare", "health policy", "vaccination", 
  "disease", "treatment", "patient care", "health education", 
  "medical research", "healthcare funding", "health insurance", 
  "preventive care", "health crisis", "nutrition", "medical ethics", 
  "health equity", "mental illness", "healthcare technology", "pandemic", 
  "health screening", "medical training", "public safety", "health regulations", 
  "mental wellness", "community health", "health services", "care homes",
  "public welfare", "health budget", "mental healthcare", "medical staffing",
  "health reforms", "social security", "child healthcare", "wellbeing",
  "health expenditure", "health data", "clinical research", "patient rights"
)

library(tm)
# get the tibble rows with NA IssueType and perform standard text preprocessing
# remove punctuation, numbers and convert to lower case
questions_tibble <- questions_tibble |> 
  filter(is.na(IssueType)) |> 
  mutate(QuestionText = str_remove_all(QuestionText, "[[:punct:]]")) |> 
  mutate(QuestionText = str_remove_all(QuestionText, "[[:digit:]]")) |> 
  mutate(QuestionText = tolower(QuestionText))

# define a function to categorise questions based on keywords
categorise_questions <- function(text, keywords_one, keywords_two) {
  if (any(str_detect(text, paste(keywords_one, collapse = "|")))) {
    if (any(str_detect(text, paste(keywords_two, collapse = "|")))) {
      return("Pending")
    } else {
      return("Economic")
    }
  } else if (any(str_detect(text, paste(keywords_two, collapse = "|")))) {
    if (any(str_detect(text, paste(keywords_one, collapse = "|")))) {
      return("Pending")
    } else {
      return("Health and Welfare")
    }
  } else {
    return("Other")
  }
}

# perform categorise function (This step may take longer to run)
questions_tibble <- questions_tibble |> 
  filter(is.na(IssueType)) |>
  mutate(IssueType = sapply(QuestionText, categorise_questions, economic_keywords, health_welfare_keywords))

# print the counts of each category
questions_tibble |> 
  group_by(IssueType) |> 
  summarise(count = n())
# Step 5, Duplicate "Pending" questions
# filter the tibble to get the questions marked as "Pending"
pending_questions <- questions_tibble |> 
  filter(IssueType == "Pending") |> 
  # duplicate each row and change the IssueType to "Economic" and "Health and Welfare"
  slice(rep(1:n(), each = 2)) |>
  mutate(IssueType = rep(c("Economic", "Health and Welfare"), length.out = nrow(pending_questions)))
  
  
  
```

### MP Information

Each question within the two tables above contains the MP Id of the MP who asked the question. We will use this information to retrieve the MP information from the MP API. We will follow the steps below to retrieve the MP information:

1.  Merge oral and written tibbles and add one column to store if each question is oral or written
2.  Create a list of all unique MP Ids from the merged tibble
3.  Loop through all MP ids to retrieve target MP info from various MP API endpoints
4.  Create a list to store MP ids that failed to return responses
5.  Recheck the failed ids

We retrieve the following information for each MP:

| Information Retrieved                                     | Members API Used                      |
|-------------------------------------------|-----------------------------|
| Displayed Name, Gender, Latest Party, Constituency, House | /api/Members/{id}                     |
| Registered Interest                                       | /api/Members/{id}/RegisteredInterests |

The failed id list is empty, meaning that all requests were successful, we store the mp tibble again in the data folder.

```{r retrieve_mp_data}
# create a list of all unique MP Ids

# read the two tibbles from the .rds files
oral_tibble <- readRDS("data/oral_tibble.rds")
written_tibble <- readRDS("data/written_tibble.rds")

# merge two tibbles and add a column to indicate if the question is oral or written
oral_tibble <- oral_tibble |> mutate(QuestionType = "oral") 
written_tibble <- written_tibble |> mutate(QuestionType = "written")
# update the column names of written tibble to match oral tibble
colnames(written_tibble) <- c("Id", "QuestionText", "TabledWhen", "AnsweringBodyId", "AnsweringBody", "AskingMemberId", "QuestionType" )
# merge the two tibbles
questions_tibble <- oral_tibble |> bind_rows(written_tibble)

# create a list of all unique MP Ids
mp_ids <- questions_tibble |> distinct(AskingMemberId) |> pull(AskingMemberId)

# create a new tibble to store MP information and set column names
mp_tibble <- tibble(
  
  AskingMemberID = NA,
  Name = NA,
  Gender = NA,
  LatestParty = NA,
  Membership = NA,
  House = NA,
  RegisteredInterest = NA
)

# create a function to retrieve a single MP's information
base_url <- "https://members-api.parliament.uk/api/Members"

failed_ids <- list()

# loop through all mp ids and retrieve MP's info, fill in the tibble
for (mp_id in mp_ids) {
  
  # a result variable to keep track of the success of current MP call
  fail <- NA
  
  # initialize all target variables to NA
  AskingMemberID <- mp_id
  Name <- NA
  Gender <- NA
  LatestParty <- NA
  Membership <- NA
  House <- NA
  RegisteredInterest <- NA
  
  # first retrieve the Displayed Name, Gender, Latest Party, Membership, House
  query_one <- sprintf("%s/%s", base_url, mp_id)
  r_one <- GET(query_one)
  if (status_code(r_one) == 200) {
    r_json <- fromJSON(rawToChar(r_one$content))
    Gender <- r_json$value$gender
    Name <- r_json$value$nameDisplayAs
    LatestParty <- r_json$value$latestParty$name
    Membership <- r_json$value$latestHouseMembership$membershipFrom
    House <- r_json$value$latestHouseMembership$house
  } else {
    print("Error")
    fail <- mp_id
  }
  
  # check_failure
  if (!is.na(fail)) {
    # add fail to failed_ids
    failed_ids <- c(failed_ids, fail)
    next
  } 
  
  # then retrieve the Registered Interest
  query_two <- sprintf("%s/%s/RegisteredInterests", base_url, mp_id)
  r_two <- GET(query_two)
  if (status_code(r_two) == 200) {
    r_json <- fromJSON(rawToChar(r_two$content))
    RegisteredInterest <- r_json$value$name
  } else {
    print("Error")
    fail <- mp_id
  }
  
  # check failure
   if (!is.na(fail)) {
     # add fail to failed_ids
    failed_ids <- c(failed_ids, fail)
    next
   } 
  
  mp_tibble <- add_row(mp_tibble, AskingMemberID = AskingMemberID, Name = Name, Gender = Gender,
                       LatestParty = LatestParty, Membership = Membership, 
                       House = House, RegisteredInterest = RegisteredInterest)
  
  # print success message followed by mp id
  print(sprintf("Success for ID%s", mp_id))
}

# save the mp tibble to data folder
saveRDS(mp_tibble, "data/mp_tibble.rds")
```

### Clean up the MP Data

Now we perform some clean up to the MP data. The tibble now has multiple rows for MPs with more than one registered interests, each row records one registered interest of a single MP. We will merge the rows of the same MP and combine the registered interests into one column. We are interested in interest category of the MPs' registered interests. Hence, when combining each row, we record the interest category number for each MP and put all numbers into a list. The category number is entailed in the name of the registered interest, either before the "." symbol or after the word "category". We will use this logic to extract the category number.

```{r clean_mp_data}

# read the mp tibble from the .rds file
mp_tibble <- readRDS("data/mp_tibble.rds")

# for each mp id, put all registered interests into a list and collapse the rows
mp_tibble <- mp_tibble |> 
  group_by(AskingMemberID) |> 
  mutate(RegisteredInterest = list(RegisteredInterest)) |> # combine all registered interests into a list
  distinct(AskingMemberID, .keep_all = TRUE) # remove duplicated rows


# print the registeredInterest of mp with askingmember id of 116
mp_tibble[mp_tibble$AskingMemberID == 116, ]$RegisteredInterest
library(stringr)
# a function to convert the list of characters to a list of numbers (retrieving category number)
convert_to_number <- function(x) {
  # initialize a list to store the numbers
  numbers <- list()
  # loop through each element in the list
  for (i in 1:length(x)) {
    # category number is either before the dot or after the word "category"
    match <- str_extract(x[i], "Category\\s*\\d+|category\\s*\\d+")
    if (!is.na(match)) {
      number <- str_extract(match, "\\d+")
      numbers <- c(numbers, number)
    } else {
      match <- str_extract(text, pattern = "\\d+\\.")
      if (!is.na(match)) {
        # Removing the dot for the final result
        number <- str_remove(match, "\\.")
        numbers <- c(numbers, number)
      }
    }
  }
  # delete duplicates of the list
  numbers <- unique(numbers)
  # return the list of numbers
  return(numbers)
}

# apply the function to each RegisteredInterest Row of the tibble
mp_tibble <- mp_tibble |> mutate(RegisteredInterest = sapply(RegisteredInterest, convert_to_number))

```

```{r read_json_files}
## Appendix: All code in this assignment

# this chunk generates the complete code appendix. 
# eval=FALSE tells R not to run (``evaluate'') the code here (it was already run before).
```
